{memer}: an R interface to the MEME Suite

### 2020-06-24 manuscript.Rmd
cleaned up split-apply-combine section

The key underlying design principle of `memer` was to create tools that easily
integrate into the existing suite of tools available in `Bioconductor` and the
broader `R` package landscape. In this way, `memer` tools can be easily integrated
into existing analysis pipelines, or used to create novel analyses.

@@

When running a MEME Suite tool using `memer`, inputs are allowed in multiple
formats providing the user additional flexibility when performing motif analysis
(see "Core Utilities" for details). Of note, most outputs of `memer` tools can
be used as inputs to other `memer` tools, allowing easy construction of
"pipelines" (for example, discovering *de-novo* motifs with DREME, then finding
the genomic positions of those motifs in target regions using FIMO).

@@

@@

For DNA and RNA sequences, the `get_sequence()` utility acts
as the entrypoint to analysis with `memer`. `get_sequence()` returns
`Biostrings::XStringSet` or `Biostrings::BStringSetList` types from range-based
data types such as `GenomicRanges` or `GenomicRangesList` to serve as
sequence input to the `memer` core functions.

@@

### Analysis using multiple groups (split-apply-combine approach)
Users can adopt a "split-apply-combine" strategy for using `memer`. `GRanges`
object acts as the centerpoint, can be split by data, fed to `get_sequence()`,
then the list will run the tools for each group automatically.

### Discriminative analysis
A key feature of several MEME Suite Tools is the ability to perform
discriminative analysis, where motifs are identified or tested for enrichement
in one group of sequences relative to another. In many instances, running tools
in discriminative mode on a set of control sequences outperforms simply using
shuffled input sequences as the control (back this up, change wording?). Therefore, 
enabling complex discriminative analysis designs was central to the design of `memer`.

When using list input to `memer` tools, names of list object members can be
passed to the `control` argument. This will result in running the tool in
discriminative mode using those sequences as the control sequences for all other
list members. By passing multiple names to `control`, the sequences will be
combined together to create a background set.

@@

### Motif Comparison (TomTom)
 - takes runDreme/meme output as input -> pipes

>>>>

The key underlying design principle of `memer` was to create tools that easily
integrate into the existing suite of tools available in `Bioconductor` and the
broader `R` package landscape. In this way, `memer` tools can be used to extend
existing analysis pipelines, or to create novel analyses.

@@

When running a MEME Suite tool using `memer`, inputs are allowed in multiple
formats providing the user additional flexibility when performing motif analysis
(see "Core Utilities" for details). Of note, most outputs of `memer` tools can
be used as inputs to other `memer` tools, allowing easy construction of
"pipelines" (for example, discovering *de-novo* motifs with DREME, then finding
the genomic positions of those motifs in target regions using FIMO).
 - unlike current MEME tools (ex tomtom -> FIMO is hard to do)

@@

**Describe each tool briefly**

**table of comparison to commandline software?**
  - input type:
  - output type:

@@

@@

A key difference between `memer` and the MEME Suite tools is that `memer`
provides tools allowing the peak-level metadata to frame subsequent analysis.
For DNA ranges (such as in ChIP-seq), the `get_sequence()` utility acts as the
entrypoint to analysis with `memer`. `get_sequence()` returns a `Biostrings`
object containing the DNA sequence of each feature in the input ranges. If
ranges are split into groups by a metadata value, `get_sequence()` will return a
list of sequences for each group. When a sequence list is used as input into
`memer` core functions, the function will run once for each group, enabling a
native "split-apply-combine" workflow without any additional user overhead
\cite{splitapplycombine, whickham}. Together, these features empower users to
use range-based metadata as the centerpoint of their analyses.

### Discriminative analysis
A powerful feature of several MEME Suite Tools is the ability to perform
discriminative analysis, where motifs are identified or tested for enrichement
in one group of sequences relative to another. In many instances, running tools
in discriminative mode on a set of control sequences outperforms simply using
shuffled input sequences as the control (back this up, change wording?). Therefore, 
enabling complex discriminative analysis designs was central to the design of `memer`.

When using list input to `memer` tools, names of list object members can be
passed to the `control` argument. This results in running the tool in
discriminative mode using those sequences as the control sequences for all other
list members. By passing multiple names to `control`, the sequences are
combined together to create a merged background set.

@@

### Motif Comparison (TomTom)
 - takes runDreme/meme output as input -> pipes
 - enables using best matched motif as input to other tools (unlike TomTom on commandline)
